---
title: "Assessment"
author: "Massih"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)

```


####  Load libraries
```{r library_load}
library(tidyverse)
library(readxl)
library(magrittr)
```
<hr></hr>

#### Read in and merge the 2010, 2011, 2012, 2013 & 2014 incidents from laser_incidents_2010-2014.xls file
##### (assume all fields marked 'UNKN' as missing values)
```{r 1_get_data}
laser10 <- read_excel("data/laser_incidents_2010-2014.xls", sheet = "Laser Report 2010")
laser11 <- read_excel("data/laser_incidents_2010-2014.xls", sheet = "Laser Report 2011")
laser12 <- read_excel("data/laser_incidents_2010-2014.xls", sheet = "Laser Report 2012")
laser13 <- read_excel("data/laser_incidents_2010-2014.xls", sheet = "Laser Report 2013")
laser14 <- read_excel("data/laser_incidents_2010-2014.xls", sheet = "Laser Report 2014")

colnames(laser11)[colnames(laser11)=="AC/ID"] <- "ACID"
colnames(laser12)[colnames(laser12)=="Aircraft ID"] <- "ACID"
colnames(laser13)[colnames(laser13)=="Aircraft ID"] <- "ACID"
colnames(laser14)[colnames(laser14)=="Aircraft ID"] <- "ACID"

laser_all <- rbind(laser10, laser11, laser12, laser13, laser14)

laser_all <- na.omit(laser_all)

laser_all$COLOR <- tolower(laser_all$COLOR)
laser_all$`Injury Reported` <- tolower(laser_all$`Injury Reported`)
laser_all$`Injury Reported` <- gsub('\\*',"",laser_all$`Injury Reported`)

laser_all$COLOR <- gsub(" ","", laser_all$COLOR)

laser_all$ALT <- gsub("[A-Z]|[a-z]",0,laser_all$ALT)
laser_all$ALT <- as.numeric(laser_all$ALT)


```
<hr></hr>

### Excluding missing/unknown values, how many distinct laser color combinations exist for 2010? 
##### Order of colors does not matter, and assume 'and', 'or', and '/' separated colors are the same
##### e.g. 'blue/green' == 'blue or green' == 'green and blue'

```{r 2_unique_color_combos_2014}
laser_all %>% 
  filter(substr(DATE,1,4) == 2010) %>% 
  distinct(laser_all, COLOR)

```
<hr></hr>

### How many incidents are identified as blue (and only blue) in the combined dataset?
```{r 3_blue_laser_incidents}
laser_all %>% 
  filter(COLOR == "blue") %>% 
  summarise(blue_sum=n())
```
<hr></hr>

### In how many different cities are laser incidents reported? Be sure to look at `CITY` and not the airport (`MAJOR CITY`).
```{r 4_cities}
laser_all %>% 
  distinct(laser_all, CITY)

```
<hr></hr>

### In what cities have *injuries* from lasers been reported?

```{r 5_injury_cities}
laser_all %>% 
  filter(`Injury Reported` != "no") %>% 
  summarise(inj_cnt = n())


```
<hr></hr>

### How many helicopters have been targeted (where 'Type A/C' is HELO)? Create a dataframe of these called 'helo'. 
```{r 6_helo}
helo <- laser_all %>% 
  filter(`TYPE A/C` == 'HELO')

summarise(helo, hel_cnt=n())

```
<hr></hr>

### In what city did laser incidents with heicopters occur most?
```{r 7_most_helo_city}
helo_city <- helo %>% 
  group_by(CITY) %>% 
  summarise(cnt=n())

helo_city[helo_city$cnt == max(helo_city$cnt)]

```
<hr></hr>

### For the combined 2010-2014 dataset, what are the maximum, minimum, and average altitudes reported?

```{r_max_min_mean_ALT}
summarise(laser_all, alt_max=max(ALT, na.rm = T), alt_min=min(ALT, na.rm = T), alt_avg=mean(ALT, na.rm = T))
```
<hr></hr>

### Create a histogram of the altitudes using ggplot2. Do the min/max/mean appear in line with  your calculations?
```{r 9_ALT_histogram}
ggplot2::ggplot(laser_all, aes(x=ALT)) + geom_histogram(bins = 70) + scale_x_continuous(limits = c(0,180000))
```
<hr></hr>

### Add a column (ALT_Category) to the dataframe based on the aircraft's altitude   
* set this to 'Low' for aircraft at or below 2000 feet; 
* 'Medium' for aircraft above 2000 feet but no higher than 8000 feet;
* 'High' for aircraft above 8000 feet.
```{r 10_create_ALT_Category}

laser_cat <- mutate(laser_all, 
       ALT_Category = case_when(
         ALT > 8000 ~ "High",
         ALT > 2000 ~ "Medium",
         ALT <= 2000 ~ "Low"
       ))
```
<hr></hr>

### Create a plot showing DATE on the x-axis and ALT on the y-axis; color the points on this plot according to the ALT_Category

```{r 11_ALT_plot}
ggplot2::ggplot(laser_cat, aes(x=DATE, y=ALT, color=ALT_Category)) + geom_point()
```


